Efficient High-Level Abstractions for Web Programming

### 2013-09-03 papers/gpce2013/gpce2013.tex
Clearer separation with ECOOP12 in the related works. More precisions in the dom creation section

%\vspace{0.1cm}
The last approach, shown in Figure \ref{fig:langs} (d), can be seen as a middle-ground between the
two previous approaches: DSLs are embedded in a host language but use a code generation process.
This approach shares the same benefits and limitations as embedded DSLs for defining language units.
However, the code generation process is specific to each DSL and gives the opportunity to perform
domain-specific optimizations. In other words deeply embedded DSLs bring domain-specific knowledge
to the compiler. Js-scala~\cite{Kossakowski12_JsDESL} is an example of deeply embedded DSL in Scala
for Web programming. It makes it possible to produce JavaScript programs from Scala code that uses
basic language concepts like arrays and control structures (\code{if} and \code{while}) as well as
mechanisms specific to the Scala compiler, like delimited continuations to handle asynchronous
computations. Though the initial authors of js-scala showed how to use the native browser APIs in an
untyped fashion, they did not show how to build high-level abstractions for Web programming
generating efficient code leveraging the browser APIs and did not show how to share such
abstractions between server and client sides (they achieved code sharing, yet on the server-side
their DSL was not compiled but interpreted). Our paper addresses these limitations.

@@

\begin{figure}[htb]
\begin{lstlisting}[language=JavaScript,label=jquery-selectors,caption=Searching elements in jQuery]
function getWords() {
  var form = $('#add-user');
  var sections = $('fieldset', form);
  return sections.map(function () {
    return $('.word', this)
  })
}
\end{lstlisting}
\end{figure}

@@

\begin{figure}[htb]
\begin{lstlisting}[language=JavaScript,caption=JavaScript DOM API,label=dom-api]
var articleUi = function (article) {
  var div = document.createElement('div');
  div.setAttribute('class', 'article');
  var span = document.createElement('span');
  var name =
    document.createTextNode(article.name + ': ');
  span.appendChild(name);
  div.appendChild(span);
  var strong = document.createElement('strong');
  var price = document.createTextNode(article.price);
  strong.appendChild(price);
  div.appendChild(strong);
  return div
};
\end{lstlisting}
\end{figure}

@@

A common task in Web applications consists in computing HTML fragments representing a part of the
page content. This task can be performed either from the server-side (to initially respond to a
request) or from the client-side (to update the current page). As an example, Listing \ref{dom-api}
defines a JavaScript function \code{articleUi} that builds a DOM tree containing an article
description, and Listing \ref{scala-xml-api} shows how one could implement a similar function on
server-side using the standard Scala XML library. The reader may notice that the client-side
and server-side APIs are very different and that the client-side API is very low-level and
inconvenient to use.

@@

Chooze~\footnote{Source code is available at
\href{http://github.com/julienrf/chooze}{http://github.com/julienrf/chooze}} is an existing
complete application for making polls. It allows users to create a poll, define the choice
alternatives, share the poll, vote and look at the results. It contains JavaScript code to handle
the dynamic behavior of the application: double-posting prevention, dynamic form update and rich
interaction with the document. The size of the whole application (server and client sides) is about one thousand of lines of code.

>>>>

%\vspace{0.1cm}
The last approach, shown in Figure \ref{fig:langs} (d), can be seen as a middle-ground between the
two previous approaches: DSLs are embedded in a host language but use a code generation process.
This approach shares the same benefits and limitations as embedded DSLs for defining language units.
However, the code generation process is specific to each DSL and gives the opportunity to perform
domain-specific optimizations. In other words deeply embedded DSLs bring domain-specific knowledge
to the compiler. Js-scala~\cite{Kossakowski12_JsDESL} is an example of deeply embedded DSL in Scala
for Web programming. It makes it possible to produce JavaScript programs from Scala code that uses
basic language concepts like arrays and control structures (\code{if} and \code{while}) as well as
mechanisms specific to the Scala compiler like delimited continuations to handle asynchronous
computations. Paper \cite{Kossakowski12_JsDESL} presented the implementation of js-scala using staging, but did not discuss any specific optimization and did not consider performance issues of this approach. In this paper, we show how js-scala has been extended to support a set of specific optimizations allowing our high-level abstractions for Web programming to be efficiently compiled on heterogeneous platforms.

@@

\begin{figure}[htb]
\begin{lstlisting}[language=JavaScript,label=jquery-selectors,caption=Searching elements using jQuery]
function getWords() {
  var form = $('#add-user');
  var sections = $('fieldset', form);
  return sections.map(function () {
    return $('.word', this)
  })
}
\end{lstlisting}
\end{figure}

@@

\begin{figure}[htb]
\begin{lstlisting}[language=JavaScript,caption=JavaScript DOM creation native API,label=dom-api]
var articleUi = function (article) {
  var div = document.createElement('div');
  div.setAttribute('class', 'article');
  var span = document.createElement('span');
  var name =
    document.createTextNode(article.name + ': ');
  span.appendChild(name);
  div.appendChild(span);
  var strong = document.createElement('strong');
  var price = document.createTextNode(article.price);
  strong.appendChild(price);
  div.appendChild(strong);
  return div
};
\end{lstlisting}
\end{figure}

@@

A common task in Web applications consists in computing HTML fragments representing a part of the
page content. This task can be performed either from the server-side (to initially respond to a
request) or from the client-side (to update the current page). As an example, Listing \ref{dom-api}
defines a JavaScript function \code{articleUi} that builds a DOM tree containing an article
description. Listing \ref{scala-xml-api} shows how one could implement a similar function on server-side using the standard Scala XML library. The reader may notice that the client-side and server-side APIs are very different and that the client-side native API is very low-level and inconvenient to use. We could use a library on client-side to get a higher level API for DOM fragment creation, but that would decrease the runtime performance. Instead, we want to define a high-level API that compiles to code as efficient as if it was written using the native APIs, on both platforms.

@@

Chooze~\footnote{Source code is available at
\href{http://github.com/julienrf/chooze}{http://github.com/julienrf/chooze}} is an existing
complete application for making polls. It allows users to create a poll, define the choice
alternatives, share the poll, vote and look at the results. It contains JavaScript code to handle
the dynamic behavior of the application: double-posting prevention, dynamic form update and rich
interaction with the document. The size of the whole application (server and client sides) is about one thousand lines of code.
